Quasi-Monte Carlo Methods: What, Why, and How?

### 2025-01-14 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin

\subsection{Discrepancies Defined by Kernels} \label{sec:kerdisc}
Let $K: [0,1)^d \times [0,1)^2 \to \mathbb{R}$ be a function satisfying the following two conditions:
\begin{subequations}
	\begin{align}
		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
		\nonumber
		\text{Postitive definiteness:} \quad & \sum_{i,j = 0}^{n-1} c_i K(\bsx_i,\bsx_j) c_j > 0 \qquad  \forall n \in \mathbb{N}, \ \bsc \ne 0, \\
		& \qquad \qquad  \ \text{distinct } \bsx_0, \ldots, \bsx_{n-1} \in [0,1)^d.
	\end{align}
\end{subequations}
Then it is known that this $K$ is the reproducing kernel for a Hilbert space, $\calH_K$ with associated inner product $\langle \cdot, \cdot \rangle_{\calH_K}$, such that
\begin{subequations}
	\begin{align}
	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d \\
	\text{Reproducing:} \quad & f(\bsx) = \langle K(\cdot,\bsx),f \rangle \qquad  \forall \bsx \in [0,1)^d, \ f \in \calH_K.
\end{align}
\end{subequations}

Having a reproducing kernel Hilbert space with a known kernel $K$ allows us to derive a rigorous error bound for $\mu - \hat{\mu}_n$.  First note that for any bounded, linear functional, $L$ on the Hilbert space $\calH_K$, there is some $\eta_L \in \calH_K$ such that
\begin{equation*}
L(f) = \langle \eta_L , f \rangle_{\calH_K} \qquad \forall f \in \calH_K.
\end{equation*}
This is guaranteed by the Riesz Representation Theorm and $\eta_L$ is called the representer of $L$.  Using the reproducing property of $K$, one may evaluate
\begin{equation*}
\eta_L(\bsx) = \langle K(\cdot, \bsx), \eta_L \rangle_{\calH_K} = L\bigl( K(\cdot,\bsx) \bigr).
\end{equation*}
From an explict expression for $\eta_L(\bsx)$ one may then calculate the squared norm of $L$ as
\begin{equation*}
	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr).
\end{equation*}

If $K$ has reasonable properties, we may use the argument above to derive our error bound.  First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.  The Riesz Representation Theorem and the Cauchy Schawarz inequality imply a tight error bound:
\begin{align}
	\nonumber
	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert \\
	& \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert  f \rVert_{\calH_K} \qquad \forall f \in \calH_K,
	\intertext{where the squared norm of the error funtional can be expressed explicitly as}
	\nonumber
	\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}^2 &=
	\bigl(\mu(\cdot\cdot) - \hat{\mu}_n(\cdot\cdot)\bigr) \Bigl(\bigl(\mu(\cdot) - \hat{\mu}_n(\cdot)\bigr)\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr) \\
	\nonumber
	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j).
\end{align}

@@

Need to finish

>>>>

\subsection{Discrepancies Defined by Kernels} \label{sec:kerdisc}
Let $K: [0,1)^d \times [0,1)^2 \to \mathbb{R}$ be a function satisfying the following two conditions:
\begin{subequations}
	\begin{align}
		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
		\nonumber
		\text{Postitive definiteness:} \quad & \sum_{i,j = 0}^{n-1} c_i K(\bsx_i,\bsx_j) c_j > 0 \qquad  \forall n \in \mathbb{N}, \ \bsc \ne 0, \\
		& \qquad \qquad  \ \text{distinct } \bsx_0, \ldots, \bsx_{n-1} \in [0,1)^d.
	\end{align}
\end{subequations}
Then it is known \cite{Aro50} that this $K$ is the reproducing kernel for a Hilbert space, $\calH_K$ with associated inner product $\langle \cdot, \cdot \rangle_{\calH_K}$, such that
\begin{subequations}
	\begin{align}
	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d \\
	\text{Reproducing:} \quad & f(\bsx) = \langle K(\cdot,\bsx),f \rangle \qquad  \forall \bsx \in [0,1)^d, \ f \in \calH_K.
\end{align}
\end{subequations}

Having a reproducing kernel Hilbert space with a known kernel $K$ allows us to derive a rigorous error bound for $\mu - \hat{\mu}_n$.  First note that for any bounded, linear functional, $L$ on the Hilbert space $\calH_K$, there is some $\eta_L \in \calH_K$ such that
\begin{equation*}
L(f) = \langle \eta_L , f \rangle_{\calH_K} \qquad \forall f \in \calH_K.
\end{equation*}
This is guaranteed by the Riesz Representation Theorm, and $\eta_L$ is called the representer of $L$.  Using the reproducing property of $K$, one may derove am explict formula for $\eta_L(\bsx)$, namely,
\begin{equation*}
\eta_L(\bsx) = \langle K(\cdot, \bsx), \eta_L \rangle_{\calH_K} = L\bigl( K(\cdot,\bsx) \bigr).
\end{equation*}
From this expression for $\eta_L(\bsx)$ one may then calculate the squared norm of the linear functional $L$ as the squared norm of its representer:
\begin{equation*}
	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr).
\end{equation*}

If $\mu(\cdot)$ is a bounded linear functional on $\calH_K$ is, we may use the argument above to derive our error bound.  First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.  The Riesz Representation Theorem and the Cauchy-Schawarz inequality imply a tight error bound:
\begin{align}
	\nonumber
	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert
	 \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert f \rVert_{\calH_K}\\
	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
	& \le \underbrace{\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}}_{\text{discrepancy}\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)}
	\, \underbrace{\inf_{c \in \RR}\lVert  f - c \rVert_{\calH_K}}_{\text{variation}(f,K)} \qquad \forall f \in \calH_K,  \label{eq:cuberrbd}
	\intertext{since $\hat{\mu}_n$ is exact for constants.  (It is assumed that $\calH$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly as}
	\nonumber
	\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}^2 &=
	\bigl(\mu(\cdot\cdot) - \hat{\mu}_n(\cdot\cdot)\bigr) \Bigl(\bigl(\mu(\cdot) - \hat{\mu}_n(\cdot)\bigr)\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr) \\
	\nonumber
	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j).
\end{align}

Error bound \eqref{eq:cuberrbd} has two factors:
\begin{itemize}
	\item the discrepancy, which is the norm of the error functional, is a measure of deficiency of the node set, $\{\bsx\}_{i=0}^{n-1}$, and
	\item the variation is a measure of the roughness of the function defining our random variable whose expectation we wish to compute.
\end{itemize}
In general, designers of qMC methods want to construct sets or sequences of nodes with  discrepancy as small as possible, either by increasing $n$, if the computational budget allows, or by better placement of the nodes.  The constructions described in Section \ref{sec:construct} are ways of ensuring better placement of the nodes, provided that the generators are chosen well.  Practitioners of qMC want to formulate $\mu$ in a way to make the variation as small as possible.  This discussed in Section \ref{sec:reformulate}.

Although error bound \eqref{eq:cuberrbd} is elegant, it leaves several questions unanswered:
\begin{itemize}
	\item The reproducing kernel, $K$, defined both the discrepancy and the variation.  This means that differnt choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
	\item The Hilbert space
\end{itemize}

@@

Need to finish and check against VS Code today
